# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

Ensure that your working directory is set to the location of activity.zip

```{r}
data <- read.csv(unz("activity.zip","activity.csv"),header=TRUE)
```
## What is mean total number of steps taken per day?

Set the scientific notation display option
```{r knitr_options}
options(scipen=10)
```

```{r}
countSteps <- aggregate(steps ~ date, data, sum)
hist(countSteps$steps, breaks = 20, main = "Total Number of Steps Taken Each Day")
mean(countSteps$steps)
median(countSteps$steps)
```

The mean total steps taken per day is `r mean(countSteps$steps)`
The median total steps taken per day is `r median(countSteps$steps)`

## What is the average daily activity pattern?
```{r}
avgsteps <- aggregate(steps ~ interval, data, mean)
plot.ts(avgsteps$steps, main="average number of steps taken in each 5-min interval")
```

The 5-min interval with the maximum average steps is `r which.max(avgsteps$steps)`

## Imputing missing values

The total number of missing values in the dataset is `r length(which(is.na(data)))`

Replace the missing values with the avg steps taken for that interval
```{r}
newdata <- data
newdata$intavg <- avgsteps[match(newdata$interval, avgsteps$interval),"steps"]
newdata$steps[is.na(newdata$steps)] <- newdata$intavg[is.na(newdata$steps)]
```
Show histogram of total no of steps taken each day
```{r}
countSteps2 <- aggregate(steps ~ date, newdata, sum)
hist(countSteps2$steps, breaks = 20, main = "Total No of Steps Taken Each Day")
```

The mean of the steps taken each day is `r mean(countSteps2$steps)` 

The difference from the original mean is `r mean(countSteps2$steps) - mean(countSteps$steps)`

The median of the staps taken each day is `r median(countSteps2$steps)`

The difference from the original median is `r median(countSteps2$steps) - median(countSteps$steps)`


## Are there differences in activity patterns between weekdays and weekends?

Add a weekday/weekend factor 

```{r}
newdata.date.f <- factor(weekdays(as.POSIXct(newdata$date)) == "Saturday" | weekdays(as.POSIXct(newdata$date)) == "Sunday", labels=c("weekday","weekend"))
```

Calculate and plot the avg daily steps grouped by the factor

```{r}
avgdailysteps <- aggregate(steps ~ interval + newdata.date.f, newdata, mean)
library("lattice")
xyplot(steps ~ interval | newdata.date.f, avgdailysteps , type = "l", layout = c(1,2,1), main = "Avg Number of Steps")
```
